---
title: "an_prep"
author: "An Bui"
date: "7/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(palmerpenguins)
library(patchwork)
library(ggimage)
library(rlang)

data("penguins")
```

We'll start with a standard plot, with counts as a function of years.

```{r}
# calculating counts per year per penguin species
penguin_counts <- penguins %>% 
  group_by(year, species) %>% 
  summarize(counts = length(species))
# look at the structure of this data frame!

# note that year has to be as.numeric
ggplot(penguin_counts, aes(x = as.numeric(year), y = counts)) +
  # geom point is the dots
  geom_point(aes(color = species), size = 5) +
  # geom_line is the lines between dots
  geom_line(aes(color = species), size = 1, alpha = 0.8) +
  
  # changing the scale of the x-axis
  scale_x_continuous(breaks = c(2007, 2008, 2009)) +
  
  # putting in colors
  scale_color_manual(values = c("darkorange", "purple", "cyan4")) + 
  
  # informative labels
  labs(x = "Year", 
       y = "Counts",
       title = "Counts per year of Adelie, Chinstrap, and Gentoo penguins",
       caption = "Data Source: Dr. Kristen Gorman, LTER Palmer Station",
       color = "Species") +
  
  # a clean theme
  theme_ipsum(grid = FALSE)
```

Now, we're going to make a plot I just learned about called a "lollipop plot." This is a similar plot to a column/bar graph, but it's especially useful when you have many columns/bars. I'm going to make them here, because I think they're fun. However, you can look at some real examples of actually useful visualizations at the [R Graph Gallery](https://www.r-graph-gallery.com/lollipop-plot.html).

```{r}
penguin_counts_island <- penguins %>% 
  group_by(year, island, species) %>% 
  summarize(counts = length(year)) %>% 
  mutate(penguin_image = case_when(
    species == "Chinstrap" ~ "chinstrap.png",
    species == "Gentoo" ~ "gentoo.png",
    species == "Adelie" ~ "adelie.png"
  )) 

island_spp <- function(which_island, which_species, which_color) {
  ggplot(penguin_counts_island %>% filter(island == {{ which_island }} & species == {{ which_species }}), aes(x = as.factor(year), y = counts)) +
    geom_segment(aes(x = as.factor(year), xend = as.factor(year), y = 0, yend = counts), color = {{ which_color }}, size = 2) +
    geom_image(aes(image = penguin_image), size = 0.1) +
    labs(title = paste(which_island, which_species, sep = " - "),
         x = "Year",
         y = "Counts") +
    theme_classic()
}
  
biscoe_adelie <- island_spp("Biscoe", "Adelie", "darkorange") +
  scale_y_continuous(limits = c(0, 20))
biscoe_adelie

biscoe_gentoo <- island_spp("Biscoe", "Gentoo", "cyan4")
biscoe_gentoo

torgersen_adelie <- island_spp("Torgersen", "Adelie", "darkorange") +
  scale_y_continuous(limits = c(0, 22))
torgersen_adelie 

dream_adelie <- island_spp("Dream", "Adelie", "darkorange") +
  scale_y_continuous(limits = c(0, 22))
dream_adelie

dream_chinstrap <- island_spp("Dream", "Chinstrap", "purple") +
  scale_y_continuous(limits = c(0, 29))
dream_chinstrap

grid <- (biscoe_adelie + biscoe_gentoo) /
        (dream_adelie + dream_chinstrap) /
        (torgersen_adelie + plot_spacer())
grid
```


