---
title: "R-Ladies Santa Barbara - Data Viz (Part 1!) - complete code handout"
author: "An Bui, Sam Csik, Juliette Verstaen"
date: "19 August 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###**Summary**

###**Outline**

**Part 0.** Load packages
**Part 1.** Juliette - TITLE HERE
**Part 2.** An - Putting your ggplot() into a function, grids
**Part 3.** Sam - Plotting summary stats using `stat_summary()`

####**Part 0.** Load packages
```{r}
library(palmerpenguins) 
library(tidyverse)
library(ggtext) 
library(hrbrthemes)
library(patchwork)
library(ggimage)
library(rlang)
```

#####################################################
####**Part 1.** Juliette
#####################################################

#####################################################
####**Part 2.** An
#####################################################

We'll start with a standard plot, with counts as a function of years.

```{r}
# calculating counts per year per penguin species
penguin_counts <- penguins %>% 
  group_by(year, species) %>% 
  summarize(counts = length(species))
# look at the structure of this data frame!

# note that year has to be as.numeric
ggplot(penguin_counts, aes(x = as.numeric(year), y = counts)) +
  # geom point is the dots
  geom_point(aes(color = species), size = 5) +
  # geom_line is the lines between dots
  geom_line(aes(color = species), size = 1, alpha = 0.8) +
  
  # changing the scale of the x-axis
  scale_x_continuous(breaks = c(2007, 2008, 2009)) +
  
  # putting in colors
  scale_color_manual(values = c("darkorange", "purple", "cyan4")) + 
  
  # informative labels
  labs(x = "Year", 
       y = "Counts",
       title = "Counts per year of Adelie, Chinstrap, and Gentoo penguins",
       caption = "Data Source: Dr. Kristen Gorman, LTER Palmer Station",
       color = "Species") +
  
  # a clean theme
  theme_ipsum()
```

Now, we're going to make a plot I just learned about called a "lollipop plot." This is a similar plot to a column/bar graph, but it's especially useful when you have many columns/bars. I'm going to make them here, because I think they're fun. However, you can look at some real examples of actually useful visualizations at the [R Graph Gallery](https://www.r-graph-gallery.com/lollipop-plot.html).

There are two packages that you can use here: the first is [ggimage](https://github.com/GuangchuangYu/ggimage), which allows you to insert images into your ggplots. The second is [patchwork](https://github.com/thomasp85/patchwork), which allows you to arrange plots in the layouts of your choosing.

First, we'll get a data frame together.
```{r}
penguin_counts_island <- penguins %>% 
  group_by(year, island, species) %>% 
  summarize(counts = length(year)) %>% 
  # adding a column of image file names for later
  mutate(penguin_image = case_when(
    species == "Chinstrap" ~ "chinstrap.png",
    species == "Gentoo" ~ "gentoo.png",
    species == "Adelie" ~ "adelie.png"
  )) 
```

Then, we're going to make a function to plot. I do this a lot when I'm exploring my data and I'm making a lot of plots that basically have the same layout but with some minor differences. 

```{r}
# the function takes 3 arguments: island, species, and color
island_spp <- function(which_island, which_species, which_color) {
  # same idea as a normal ggplot function, except now you're making your own!
  # step 1. filter the data frame to only include the island and species of interest
  ggplot(penguin_counts_island %>% filter(island == {{ which_island }} & species == {{ which_species }}), 
         # set the x and y axes
         aes(x = as.factor(year), y = counts)) +
    
    # the segment is the "stick" of the lollipop - note the start and end
    geom_segment(aes(x = as.factor(year), xend = as.factor(year), y = 0, yend = counts), color = {{ which_color }}, size = 2) +
    
    # the image is the candy part!
    geom_image(aes(image = penguin_image), size = 0.1) +
    
    # labels
    labs(title = paste(which_island, which_species, sep = " - "),
         x = "Year",
         y = "Counts") +
    
    # theme
    theme_classic()
}
```

Now, we're ready to plot!

```{r}
biscoe_adelie <- island_spp("Biscoe", "Adelie", "darkorange") +
  scale_y_continuous(limits = c(0, 20))
biscoe_adelie

biscoe_gentoo <- island_spp("Biscoe", "Gentoo", "cyan4")
biscoe_gentoo

torgersen_adelie <- island_spp("Torgersen", "Adelie", "darkorange") +
  scale_y_continuous(limits = c(0, 22))
torgersen_adelie 

dream_adelie <- island_spp("Dream", "Adelie", "darkorange") +
  scale_y_continuous(limits = c(0, 22))
dream_adelie

dream_chinstrap <- island_spp("Dream", "Chinstrap", "purple") +
  scale_y_continuous(limits = c(0, 29))
dream_chinstrap

# now put everything together with patchwork!
grid <- (biscoe_adelie + biscoe_gentoo) /
        (dream_adelie + dream_chinstrap) /
        (torgersen_adelie + plot_spacer()) +
  # add plot labels
  plot_annotation(tag_levels = "a", tag_suffix = ")")

grid
```

#####################################################
####**Part 3.** Sam
#####################################################

Here, we'll walk through plotting a combination of raw data and summary stats together on the same figure. There are two general approaches:

(1) calculate summary stats and save to a new data frame; plot raw data from original data frame and layer on summary stats from the new data frame -- this tends to require **more** code

(2) use the `stat_summary()` function within ggplot to simultaneously calculate and plot summary stats 

We'll recreate the same figure using both approaches.

*Apporach 1* Longer way: Summarize data for plotting, then construct figure 
```{r}
##############################
# summarize body mass data (mean, max, min)
##############################

bm_summary <- penguins %>% 
  drop_na() %>%
  group_by(species, sex) %>% 
  summarize(
    mean = mean(body_mass_g),
    sd = sd(body_mass_g)
  )

##############################
# plot
##############################

ggplot() +
  
  # plot raw data
  geom_point(data = na.omit(penguins), aes(x = species, y = body_mass_g, shape = sex), 
             alpha = 0.2, 
             position = position_jitterdodge(dodge.width = 0.8)) +
  
  # layer on summary stats
  geom_errorbar(data = bm_summary, aes(x = species, ymin = mean - sd, ymax = mean + sd, color = species, group = sex), 
                width = 0.1, size = 1,
                position = position_dodge(width = 0.8)) +
  geom_point(data = bm_summary, aes(x = species, y = mean, shape = sex, color = species),
             size = 4,
             position = position_dodge(width = 0.8)) +
  
  # manually adjust colors and shapes (gotta reflect what's up )
  scale_color_manual(values = c("darkorange", "purple", "cyan4"), guide = FALSE) + # set `guide = FALSE` to omit from legend
  scale_shape_manual(values = c(15, 16), name = "Sex") + # set `name =` to change legend title
  
  # add nicer axis labels + caption
  labs(x = "Penguin Species", 
       y = "Body Mass (g)",
       title = "Body weights of female vs. male <b style='color:darkorange'>adelie,</b> <b style='color:purple'>chinstrap,</b> and <b style='color:cyan4'>gentoo</b> penguins",
       subtitle = "Colored points represent mean body mass (± SD)",
       caption = "Data Source: Dr. Kristen Gorman, LTER Palmer Station") +
  
  # now fun with themes!
  theme_ipsum(grid = FALSE) +
  theme(
        # text
        plot.title = element_markdown(hjust = 0, size = 13),
        # axis.text = element_text(color = "black", size = 10), 
        # axis.title = element_text(size = 14), 
        plot.caption = element_text(size = 7, hjust = 1, color = "gray", face = "italic"),
        # plot
        panel.border = element_rect(colour = "black", size = 0.7, fill = NA), 
        # legend
        legend.position = c(0.12, 0.85), # (right/left, top/bottom)
        legend.box.background = element_rect(color = "black")
        ) 
```

*Approach 2* Shortcut using `stat_summary()`; benefits: (a) no need to create a second df with summary stats, (b) only need to set overall aesthetics once in the ggplot() function
```{r}
##############################
# plot
##############################

# plot raw data
ggplot(data = na.omit(penguins), aes(x = species, y = ___, shape = ___)) + # choose a different variable, or stick with body_mass_g :)
  geom_point(alpha = ___, position = position_jitterdodge(dodge.width = ___)) + 
  
  # calculate and layer on summary stats (mean +/- sd)
  stat_summary(mapping = aes(color = ___),
               fun = "mean", geom = "___", size = ___,
               position = position_dodge(width = ___)) +
  stat_summary(mapping = aes(color = species),
               fun = "mean", geom = "errorbar", size = 1, width = 0.2,
               fun.max = ___,
               fun.min = ___,
               position = position_dodge(width = ___)) +
  
  # manually adjust colors and shapes
  scale_color_manual(values = c("___", "___", "___"), guide = FALSE) + # set `guide = FALSE` to omit from legend
  scale_shape_manual(values = c(___, ___), name = "Sex") + # set `name =` to change legend title
  
  # add nicer axis labels + title + caption
  labs(x = "Penguin Species", 
       y = "___",
       title = "Body weights of female vs. male adelie, chinstrap, and gentoo penguins",
       subtitle = "Colored points represent mean body mass (± SD)",
       caption = "Data Source: Dr. Kristen Gorman, LTER Palmer Station") +
  
  ########################
  # theme option 1 (basic ggplot)
  ########################
  theme___*__() + # try out different built-in themes e.g theme_minimal(), theme_bw(), etc
  theme(
        # text
        plot.title = element_text(hjust = ___, size = ___),
        axis.text = element_text(color = "___", size = ___),
        axis.title = element_text(color = "___", size = ___),
        plot.caption = element_text(size = ___, hjust = ___, color = "___", face = "___"),

        # plot
        panel.border = element_rect(color = "___", size = ___, fill = NA),

        # legend
        legend.position = c(___, ___), # c(right/left, top/bottom)
        legend.box.background = element_rect(color = "___"))

  ########################
  # theme option 2 (a little bit fancier)
  ########################
  # theme_ipsum(grid = ___) + # from hrbrthemes package
  # theme(
  #       # text
  #       plot.title = element_markdown(hjust = ___ size = ___),
  #       plot.caption = element_text(size = ___, hjust = ___, color = "___", face = "___"),
  # 
  #       # plot
  #       panel.border = element_rect(colour = "___", size = ___, fill = NA),
  # 
  #       # legend
  #       legend.position = c(0___, ___), # c(right/left, top/bottom)
  #       legend.box.background = element_rect(color = "___"))

# title for use with ggtext() : "Body weights of female vs. male <b style='color:___'>adelie,</b> <b style='color:___'>chinstrap,</b> and <b style='color:___'>gentoo</b> penguins"

########################
# Optional: save your plot as a .jpg, .pdf, etc. using ggsave()
########################

# ggsave(plot = SC_simple, here::here("plots", "SC_simple.jpg"), width = 7, height = 7)
```

