---
title: "R-Ladies Santa Barbara - Data Viz (Part 1!) - complete code handout"
author: "An Bui, Sam Csik, Juliette Verstaen"
date: "19 August 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###**Summary**

We'll be walking through the components necessary for constructing a basic ggplot, and also delve into some cool features for elevating the aesthetics of your plots.

*NOTE: Normally, you should take a critical look at your data, determine why NAs exist and if they might bias your results...we're skipping that part here and just omitting any NAs for the sake of time. But for when you do explore NAs, we suggest checking out the `naniar` package*

###**Outline**

**Part 0.** Load packages
**Part 1.** Juliette - Basic ggplot(), putting the legend in the title, and themes
**Part 2.** An - Putting your ggplot() into a function, grids
**Part 3.** Sam - Plotting summary stats using `stat_summary()`

####**Part 0.** Load packages
```{r}
library(palmerpenguins) 
library(tidyverse)
library(ggtext) 
library(hrbrthemes)
library(patchwork)
library(ggimage)
library(rlang)
```

#####################################################
####**Part 1.** Juliette
#####################################################

We're going to create a standard scatter ggplot and show the 3 main things I do every time I create a plot:
1. Choose the type of plot you want and the aesthetics needed to make it happen
2. Labels
3. Make it look pretty! - for me this means: what the plot theme is, messing with things in the theme(), and playing with colors

I always do this part last because I spend forever on it, and I want to get the meat of the plot right first

Then I’m going to show you how to represent your legend (if colors) in the title and some plot themes from the package ‘hrbrthemes’


```{r create scatter}

ggplot(na.omit(penguins)) +
   
   ## 1. Choose your plot and aesthetics
   geom_point(aes(x = body_mass_g, y = flipper_length_mm, color = species, shape = island), size = 2) +
   
   ## 2. Label all the things
   labs(x = "Body Mass (g)",
        y = "Flipper Length (mm)",
        title = "Exploring Palmer Station Penguins",
        subtitle = "Dimensions and Attributes",
        caption = "Data Source: Dr. Kristen Gorman LTER Palmer Station",
        shape = "Island",
        color = "Species") +
   
   ## 3a. Built in theme
   theme_minimal() + 
   
   ## 3b. Editing the labels aesthetics
   theme(plot.title = element_text(color = "black", size = 18, face = "bold", hjust = 0.5),
         plot.subtitle = element_text(color = "gray41", size = 14, face = "bold", hjust = 0.5),
         plot.caption = element_text(color = "gray65", face = "italic"),
         axis.title.x = element_text(size = 15),
         axis.title.y = element_text(size = 15)) +
   
   ## 3c. Change colors
   scale_colour_manual(values =  c("darkorange", "purple", "cyan4")) +
   
   ## 3d. Faceting
   #facet_wrap(vars(sex)) 
   facet_grid(cols = vars(sex), rows = vars(island))

```

Now we are going to make the same plot incorporating the legend into the title and a pretty new theme!!

```{r create prettier scatter}

ggplot(na.omit(penguins)) +
 
   ## 1. Choose your plot and aesthetics
   geom_point(aes(x = body_mass_g, y = flipper_length_mm, color = species)) +
   
   ## 2. Label all the things
   labs(x = "Body Mass (g)",
        y = "Flipper Length (mm)",
        
        ## use html/markdown style to code the colors into the title
        ## title : "Dimensions for adelie, chinstrap, and gentoos penguins"
        ## We need to tell ggplot that:  html format text is coming up, what we want that text to look like, what the text says, and that html format text is over
        title = "Dimensions for <b style='color:darkorange'>adelie,</b> <b style='color:purple'>chinstrap,</b> and <b style='color:cyan4'>gentoos</b> penguins",
        
        caption = "Data Source: Dr. Kristen Gorman LTER Palmer Station \nCreated for RLadies SB workshop") +
   
   ## 3a. Choose an 'hrbrthemes' theme
   theme_ipsum_rc() +
   
   ## 3b. Editing the labels aesthetics
   theme(plot.title = element_markdown(hjust = 0.5, size = 20),
         plot.caption = element_text(color = "gray65", face = "italic", size = 8),
         legend.position = "none") +
   
   ## 3c. Change colors
   scale_colour_manual(values = col_pal) +
   
   ## 3d. Faceting
   facet_grid(cols = vars(sex), rows = vars(island)) 

```





#####################################################
####**Part 2.** An
#####################################################

We'll start with a standard plot, with counts as a function of years.

```{r}
# calculating counts per year per penguin species
penguin_counts <- penguins %>% 
  group_by(year, species) %>% 
  summarize(counts = length(species))
# look at the structure of this data frame!

# note that year has to be as.numeric
ggplot(penguin_counts, aes(x = as.numeric(year), y = counts)) +
  # geom point is the dots
  geom_point(aes(color = species), size = 5) +
  # geom_line is the lines between dots
  geom_line(aes(color = species), size = 1, alpha = 0.8) +
  
  # changing the scale of the x-axis
  scale_x_continuous(breaks = c(2007, 2008, 2009)) +
  
  # putting in colors
  scale_color_manual(values = c("darkorange", "purple", "cyan4")) + 
  
  # informative labels
  labs(x = "Year", 
       y = "Counts",
       title = "Counts per year of Adelie, Chinstrap, and Gentoo penguins",
       caption = "Data Source: Dr. Kristen Gorman, LTER Palmer Station",
       color = "Species") +
  
  # a clean theme
  theme_ipsum()
```

Now, we're going to make a plot I just learned about called a "lollipop plot." This is a similar plot to a column/bar graph, but it's especially useful when you have many columns/bars. I'm going to make them here, because I think they're fun. However, you can look at some real examples of actually useful visualizations at the [R Graph Gallery](https://www.r-graph-gallery.com/lollipop-plot.html).

There are two packages that you can use here: the first is [ggimage](https://github.com/GuangchuangYu/ggimage), which allows you to insert images into your ggplots. The second is [patchwork](https://github.com/thomasp85/patchwork), which allows you to arrange plots in the layouts of your choosing.

First, we'll get a data frame together.
```{r}
penguin_counts_island <- penguins %>% 
  group_by(year, island, species) %>% 
  summarize(counts = length(year)) %>% 
  # adding a column of image file names for later
  mutate(penguin_image = case_when(
    species == "Chinstrap" ~ "chinstrap.png",
    species == "Gentoo" ~ "gentoo.png",
    species == "Adelie" ~ "adelie.png"
  )) 
```

Then, we're going to make a function to plot. I do this a lot when I'm exploring my data and I'm making a lot of plots that basically have the same layout but with some minor differences. 

```{r}
# the function takes 3 arguments: island, species, and color
island_spp <- function(which_island, which_species, which_color) {
  # same idea as a normal ggplot function, except now you're making your own!
  # step 1. filter the data frame to only include the island and species of interest
  ggplot(penguin_counts_island %>% filter(island == {{ which_island }} & species == {{ which_species }}), 
         # set the x and y axes
         aes(x = as.factor(year), y = counts)) +
    
    # the segment is the "stick" of the lollipop - note the start and end
    geom_segment(aes(x = as.factor(year), xend = as.factor(year), y = 0, yend = counts), color = {{ which_color }}, size = 2) +
    
    # the image is the candy part!
    geom_image(aes(image = penguin_image), size = 0.1) +
    
    # labels
    labs(title = paste(which_island, which_species, sep = " - "),
         x = "Year",
         y = "Counts") +
    
    # theme
    theme_classic()
}
```

Now, we're ready to plot!

```{r}
biscoe_adelie <- island_spp("Biscoe", "Adelie", "darkorange") +
  scale_y_continuous(limits = c(0, 20))
biscoe_adelie

biscoe_gentoo <- island_spp("Biscoe", "Gentoo", "cyan4")
biscoe_gentoo

torgersen_adelie <- island_spp("Torgersen", "Adelie", "darkorange") +
  scale_y_continuous(limits = c(0, 22))
torgersen_adelie 

dream_adelie <- island_spp("Dream", "Adelie", "darkorange") +
  scale_y_continuous(limits = c(0, 22))
dream_adelie

dream_chinstrap <- island_spp("Dream", "Chinstrap", "purple") +
  scale_y_continuous(limits = c(0, 29))
dream_chinstrap

# now put everything together with patchwork!
grid <- (biscoe_adelie + biscoe_gentoo) /
        (dream_adelie + dream_chinstrap) /
        (torgersen_adelie + plot_spacer()) +
  # add plot labels
  plot_annotation(tag_levels = "a", tag_suffix = ")")

grid
```

#####################################################
####**Part 3.** Sam
#####################################################

Here, we'll walk through plotting a combination of raw data and summary stats together on the same figure. There are two general approaches:

(1) calculate summary stats and save to a new data frame; plot raw data from original data frame and layer on summary stats from the new data frame -- this tends to require **more** code

(2) use the `stat_summary()` function within ggplot to simultaneously calculate and plot summary stats 

We'll recreate the same figure using both approaches.

*Apporach 1* Longer way: Summarize data for plotting, then construct figure 
```{r}
##############################
# summarize body mass data (mean, max, min)
##############################

bm_summary <- penguins %>% 
  drop_na() %>%
  group_by(species, sex) %>% 
  summarize(
    mean = mean(body_mass_g),
    sd = sd(body_mass_g)
  )

##############################
# plot
##############################

ggplot() +
  
  # plot raw data
  geom_point(data = na.omit(penguins), aes(x = species, y = body_mass_g, shape = sex), 
             alpha = 0.2, 
             position = position_jitterdodge(dodge.width = 0.8)) +
  
  # layer on summary stats
  geom_errorbar(data = bm_summary, aes(x = species, ymin = mean - sd, ymax = mean + sd, color = species, group = sex), 
                width = 0.1, size = 1,
                position = position_dodge(width = 0.8)) +
  geom_point(data = bm_summary, aes(x = species, y = mean, shape = sex, color = species),
             size = 4,
             position = position_dodge(width = 0.8)) +
  
  # manually adjust colors and shapes (gotta reflect what's up )
  scale_color_manual(values = c("darkorange", "purple", "cyan4"), guide = FALSE) + # set `guide = FALSE` to omit from legend
  scale_shape_manual(values = c(15, 16), name = "Sex") + # set `name =` to change legend title
  
  # add nicer axis labels + caption
  labs(x = "Penguin Species", 
       y = "Body Mass (g)",
       title = "Body mass of female vs. male <b style='color:darkorange'>adelie,</b> <b style='color:purple'>chinstrap,</b> and <b style='color:cyan4'>gentoo</b> penguins",
       subtitle = "Colored points represent mean body mass (± SD)",
       caption = "Data Source: Dr. Kristen Gorman, LTER Palmer Station") +
  
  # now fun with themes!
  theme_ipsum(grid = FALSE) +
  theme(
        # text
        plot.title = element_markdown(hjust = 0, size = 13),
        # axis.text = element_text(color = "black", size = 10), 
        # axis.title = element_text(size = 14), 
        plot.caption = element_text(size = 7, hjust = 1, color = "gray", face = "italic"),
        # plot
        panel.border = element_rect(colour = "black", size = 0.7, fill = NA), 
        # legend
        legend.position = c(0.12, 0.85), # (right/left, top/bottom)
        legend.box.background = element_rect(color = "black")
        ) 
```

*Approach 2* Shortcut using `stat_summary()`; benefits: (a) no need to create a second df with summary stats, (b) only need to set overall aesthetics once in the ggplot() function
```{r}
##############################
# plot
##############################

# NOTE: you can sub "pd" anywhere you see `position_jitterdodge(dodge.width = 0.8)` if you prefer
# pd <- position_jitterdoge(dodge.width = 0.8)

# plot raw data
ggplot(data = na.omit(penguins), aes(x = species, y = body_mass_g, shape = sex)) +
  geom_point(alpha = 0.2, position = position_jitterdodge(dodge.width = 0.8)) + 
  
  # calculate and layer on summary stats (mean +/- sd)
  stat_summary(mapping = aes(color = species),
               fun = "mean", geom = "point", size = 4,
               position = position_dodge(width = 0.8)) +
  stat_summary(mapping = aes(color = species),
               fun = "mean", geom = "errorbar", size = 1, width = 0.2,
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               position = position_dodge(width = 0.8)) +
  
  # manually adjust colors and shapes
  scale_color_manual(values = c("darkorange", "purple", "cyan4"), guide = FALSE) + # set `guide = FALSE` to omit from legend
  scale_shape_manual(values = c(15, 16), name = "Sex") + # set `name =` to change legend title
  
  # add nicer axis labels + title + caption
  labs(x = "Penguin Species", 
       y = "Body Mass (g)",
       title = "Body mass of female vs. male adelie, chinstrap, and gentoo penguins",
       subtitle = "Colored points represent mean body mass (± SD)",
       caption = "Data Source: Dr. Kristen Gorman, LTER Palmer Station") +
  
  ########################
  # theme option 1 (basic ggplot)
  ########################
  theme_classic() + # built-in theme
  theme(
        # text
        plot.title = element_text(hjust = 0, size = 14),
        axis.text = element_text(color = "black", size = 10),
        axis.title = element_text(color = "black", size = 14),
        plot.caption = element_text(size = 7, hjust = 1, color = "gray", face = "italic"),

        # plot
        panel.border = element_rect(color = "black", size = 0.7, fill = NA),

        # legend
        legend.position = c(0.12, 0.85), # c(right/left, top/bottom)
        legend.box.background = element_rect(color = "black"))

  ########################
  # theme option 2 (a little bit fancier)
  ########################
  # theme_ipsum(grid = FALSE) + # from hrbrthemes package
  # theme(
  #       # text
  #       plot.title = element_markdown(hjust = 0, size = 13),
  #       plot.caption = element_text(size = 7, hjust = 1, color = "gray", face = "italic"),
  # 
  #       # plot
  #       panel.border = element_rect(colour = "black", size = 0.7, fill = NA),
  # 
  #       # legend
  #       legend.position = c(0.12, 0.85), # c(right/left, top/bottom)
  #       legend.box.background = element_rect(color = "black"))

# title for use with ggtext() : "Body mass of female vs. male <b style='color:darkorange'>adelie,</b> <b style='color:purple'>chinstrap,</b> and <b style='color:cyan4'>gentoo</b> penguins"

########################
# Optional: save your plot as a .jpg, .pdf, etc. using ggsave()
########################

# ggsave(plot = SC_simple, here::here("plots", "SC_simple.jpg"), width = 7, height = 7)
```

