---
title: "R-Ladies Santa Barbara - Data Viz (Part 1!) - fill-in worksheet"
author: "An Bui, Sam Csik, Juliette Verstaen"
date: "19 August 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###**Summary**

###**Outline**

**Part 0.** Load packages
**Part 1.** Juliette - TITLE HERE
**Part 2.** An - TITLE HERE
**Part 3.** Sam - Plotting summary stats using `stat_summary()`

####**Part 0.** Load packages
```{r}
library(palmerpenguins) 
library(tidyverse)
library(ggtext) 
library(hrbrthemes)
library(pathwork)
```

#####################################################
####**Part 1.** Juliette
#####################################################

We're going to create a standard scatter ggplot and show the 3 main things I do every time I create a plot:
1. Choose the type of plot you want and the aesthetics needed to make it happen
2. Labels
3. Make it look pretty! - for me this means: what the plot theme is, messing with things in the theme(), and playing with colors

I always do this part last because I spend forever on it, and I want to get the meat of the plot right first

Then I’m going to show you how to represent your legend (if colors) in the title and some plot themes from the package ‘hrbrthemes’


```{r create scatter}

ggplot(na.omit(penguins)) +
   
   ## 1. Choose your plot and aesthetics
   geom_point(aes(x = body_mass_g, y = flipper_length_mm, color = species, shape = island), size = 2) +
   
   ## 2. Label all the things
   labs(x = "Body Mass (g)",
        y = "Flipper Length (mm)",
        title = "Exploring Palmer Station Penguins",
        subtitle = "Dimensions and Attributes",
        caption = "Data Source: Dr. Kristen Gorman LTER Palmer Station",
        shape = "Island",
        color = "Species") +
   
   ## 3a. Built in theme
   theme_minimal() + 
   
   ## 3b. Editing the labels aesthetics
   theme(plot.title = element_text(color = "black", size = 18, face = "bold", hjust = 0.5),
         plot.subtitle = element_text(color = "gray41", size = 14, face = "bold", hjust = 0.5),
         plot.caption = element_text(color = "gray65", face = "italic"),
         axis.title.x = element_text(size = 15),
         axis.title.y = element_text(size = 15)) +
   
   ## 3c. Change colors
   scale_colour_manual(values =  c("darkorange", "purple", "cyan4")) +
   
   ## 3d. Faceting
   #facet_wrap(vars(sex)) 
   facet_grid(cols = vars(sex), rows = vars(island))

```

Now we are going to make the same plot incorporating the legend into the title and a pretty new theme!!

```{r create prettier scatter}

ggplot(na.omit(penguins)) +
 
   ## 1. Choose your plot and aesthetics
   geom_point(aes(x = ____, y = ____, color = ____)) +
   
   ## 2. Label all the things
   labs(x = "____",
        y = "____",
        
        ## use html/markdown style to code the colors into the title
        ## title : "Dimensions for adelie, chinstrap, and gentoos penguins"
        ## We need to tell ggplot that:  html format text is coming up, what we want that text to look like, what the text says, and that html format text is over
        title = "Dimensions for <b style='color:____'>____,</b> <b style='color:____'>____,</b> and <b style='color:____'>____</b> penguins",
        
        caption = "Data Source: Dr. Kristen Gorman LTER Palmer Station \nCreated for RLadies SB workshop") +
   
   ## 3a. Choose an 'hrbrthemes' theme; play around with them!
   theme_ipsum_rc() +
   
   ## 3b. Editing the labels aesthetics
   theme(plot.title = element_markdown(hjust = 0.5, size = 20),
         plot.caption = element_text(color = "gray65", face = "italic", size = 8),
         legend.position = "none") +
   
   ## 3c. Change colors
   scale_colour_manual(values =  c("darkorange", "purple", "cyan4")) +
   
   ## 3d. Faceting
   facet_grid(cols = vars(sex), rows = vars(island)) 

```



#####################################################
####**Part 2.** An
#####################################################

#####################################################
####**Part 3.** Sam
#####################################################

Here, we'll walk through plotting a combination of raw data and summary stats together on the same figure. There are two general approaches:

(1) calculate summary stats and save to a new data frame; plot raw data from original data frame and layer on summary stats from the new data frame -- this tends to require **more** code

(2) use the `stat_summary()` function within ggplot to simultaneously calculate and plot summary stats 

We'll recreate the same figure using both approaches.

*Apporach 1* Longer way: Summarize data for plotting, then construct figure
```{r}
##############################
# summarize body mass data (mean, max, min)
##############################

bm_summary <- penguins %>% 
  drop_na() %>%
  group_by(species, sex) %>% 
  summarize(
    mean = mean(body_mass_g),
    sd = sd(body_mass_g)
  )

##############################
# plot
##############################

ggplot() +
  
  # plot raw data
  geom_point(data = na.omit(penguins), aes(x = species, y = body_mass_g, shape = sex), 
             alpha = 0.2, 
             position = position_jitterdodge(dodge.width = 0.8)) +
  
  # layer on summary stats
  geom_errorbar(data = bm_summary, aes(x = species, ymin = mean - sd, ymax = mean + sd, color = species, group = sex), 
                width = 0.1, size = 1,
                position = position_dodge(width = 0.8)) +
  geom_point(data = bm_summary, aes(x = species, y = mean, shape = sex, color = species),
             size = 4,
             position = position_dodge(width = 0.8)) +
  
  # manually adjust colors and shapes (gotta reflect what's up )
  scale_color_manual(values = c("darkorange", "purple", "cyan4"), guide = FALSE) + # set `guide = FALSE` to omit from legend
  scale_shape_manual(values = c(15, 16), name = "Sex") + # set `name =` to change legend title
  
  # add nicer axis labels + caption
  labs(x = "Penguin Species", 
       y = "Body Mass (g)",
       title = "Body weights of female vs. male <b style='color:darkorange'>adelie,</b> <b style='color:purple'>chinstrap,</b> and <b style='color:cyan4'>gentoo</b> penguins",
       subtitle = "Colored points represent mean body mass (± SD)",
       caption = "Data Source: Dr. Kristen Gorman, LTER Palmer Station") +
  
  # now fun with themes!
  theme_ipsum(grid = FALSE) +
  theme(
        # text
        plot.title = element_markdown(hjust = 0, size = 13),
        # axis.text = element_text(color = "black", size = 10), 
        # axis.title = element_text(size = 14), 
        plot.caption = element_text(size = 7, hjust = 1, color = "gray", face = "italic"),
        # plot
        panel.border = element_rect(colour = "black", size = 0.7, fill = NA), 
        # legend
        legend.position = c(0.12, 0.85), # (right/left, top/bottom)
        legend.box.background = element_rect(color = "black")
        ) 
```

*Approach 2* Shortcut using `stat_summary()`; benefits: (a) no need to create a second df with summary stats, (b) only need to set overall aesthetics once in the ggplot() function
```{r}
##############################
# plot
##############################

# plot raw data
ggplot(data = na.omit(penguins), aes(x = species, y = ___, shape = ___)) + # choose a different variable, or stick with body_mass_g :)
  geom_point(alpha = ___, position = position_jitterdodge(dodge.width = ___)) + 
  
  # calculate and layer on summary stats (mean +/- sd)
  stat_summary(mapping = aes(color = ___),
               fun = "mean", geom = "___", size = ___,
               position = position_dodge(width = ___)) +
  stat_summary(mapping = aes(color = species),
               fun = "mean", geom = "errorbar", size = 1, width = 0.2,
               fun.max = ___,
               fun.min = ___,
               position = position_dodge(width = ___)) +
  
  # manually adjust colors and shapes
  scale_color_manual(values = c("___", "___", "___"), guide = FALSE) + # set `guide = FALSE` to omit from legend
  scale_shape_manual(values = c(___, ___), name = "Sex") + # set `name =` to change legend title
  
  # add nicer axis labels + title + caption
  labs(x = "Penguin Species", 
       y = "___",
       title = "Body weights of female vs. male adelie, chinstrap, and gentoo penguins",
       subtitle = "Colored points represent mean body mass (± SD)",
       caption = "Data Source: Dr. Kristen Gorman, LTER Palmer Station") +
  
  ########################
  # theme option 1 (basic ggplot)
  ########################
  theme___*__() + # try out different built-in themes e.g theme_minimal(), theme_bw(), etc
  theme(
        # text
        plot.title = element_text(hjust = ___, size = ___),
        axis.text = element_text(color = "___", size = ___),
        axis.title = element_text(color = "___", size = ___),
        plot.caption = element_text(size = ___, hjust = ___, color = "___", face = "___"),

        # plot
        panel.border = element_rect(color = "___", size = ___, fill = NA),

        # legend
        legend.position = c(___, ___), # c(right/left, top/bottom)
        legend.box.background = element_rect(color = "___"))

  ########################
  # theme option 2 (a little bit fancier)
  ########################
  # theme_ipsum(grid = ___) + # from hrbrthemes package
  # theme(
  #       # text
  #       plot.title = element_markdown(hjust = ___ size = ___),
  #       plot.caption = element_text(size = ___, hjust = ___, color = "___", face = "___"),
  # 
  #       # plot
  #       panel.border = element_rect(colour = "___", size = ___, fill = NA),
  # 
  #       # legend
  #       legend.position = c(0___, ___), # c(right/left, top/bottom)
  #       legend.box.background = element_rect(color = "___"))

# title for use with ggtext() : "Body weights of female vs. male <b style='color:___'>adelie,</b> <b style='color:___'>chinstrap,</b> and <b style='color:___'>gentoo</b> penguins"

########################
# Optional: save your plot as a .jpg, .pdf, etc. using ggsave()
########################

# ggsave(plot = SC_simple, here::here("plots", "SC_simple.jpg"), width = 7, height = 7)
```