---
title: "R-Ladies Santa Barbara - Data Viz (Part 1!) - complete code handout"
author: "An Bui, Sam Csik, Juliette Verstaen"
date: "19 August 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###**Summary**

###**Outline**

**Part 0.** Load packages
**Part 1.** Juliette - TITLE HERE
**Part 2.** An - TITLE HERE
**Part 3.** Sam - Plotting summary stats using `stat_summary()`

####**Part 0.** Load packages
```{r}
library(palmerpenguins) 
library(tidyverse)
library(ggtext) 
library(hrbrthemes)
library(patchwork)
```

#####################################################
####**Part 1.** Juliette
#####################################################

#####################################################
####**Part 2.** An
#####################################################

#####################################################
####**Part 3.** Sam
#####################################################

Here, we'll walk through plotting a combination of raw data and summary stats together on the same figure. There are two general approaches:

(1) calculate summary stats and save to a new data frame; plot raw data from original data frame and layer on summary stats from the new data frame -- this tends to require **more** code

(2) use the `stat_summary()` function within ggplot to simultaneously calculate and plot summary stats 

We'll recreate the same figure using both approaches.

*Apporach 1* Longer way: Summarize data for plotting, then construct figure 
```{r}
##############################
# summarize body mass data (mean, max, min)
##############################

bm_summary <- penguins %>% 
  drop_na() %>%
  group_by(species, sex) %>% 
  summarize(
    mean = mean(body_mass_g),
    sd = sd(body_mass_g)
  )

##############################
# plot
##############################

ggplot() +
  
  # plot raw data
  geom_point(data = na.omit(penguins), aes(x = species, y = body_mass_g, shape = sex), 
             alpha = 0.2, 
             position = position_jitterdodge(dodge.width = 0.8)) +
  
  # layer on summary stats
  geom_errorbar(data = bm_summary, aes(x = species, ymin = mean - sd, ymax = mean + sd, color = species, group = sex), 
                width = 0.1, size = 1,
                position = position_dodge(width = 0.8)) +
  geom_point(data = bm_summary, aes(x = species, y = mean, shape = sex, color = species),
             size = 4,
             position = position_dodge(width = 0.8)) +
  
  # manually adjust colors and shapes (gotta reflect what's up )
  scale_color_manual(values = c("darkorange", "purple", "cyan4"), guide = FALSE) + # set `guide = FALSE` to omit from legend
  scale_shape_manual(values = c(15, 16), name = "Sex") + # set `name =` to change legend title
  
  # add nicer axis labels + caption
  labs(x = "Penguin Species", 
       y = "Body Mass (g)",
       title = "Body weights of female vs. male <b style='color:darkorange'>adelie,</b> <b style='color:purple'>chinstrap,</b> and <b style='color:cyan4'>gentoo</b> penguins",
       subtitle = "Colored points represent mean body mass (± SD)",
       caption = "Data Source: Dr. Kristen Gorman, LTER Palmer Station") +
  
  # now fun with themes!
  theme_ipsum(grid = FALSE) +
  theme(
        # text
        plot.title = element_markdown(hjust = 0, size = 13),
        # axis.text = element_text(color = "black", size = 10), 
        # axis.title = element_text(size = 14), 
        plot.caption = element_text(size = 7, hjust = 1, color = "gray", face = "italic"),
        # plot
        panel.border = element_rect(colour = "black", size = 0.7, fill = NA), 
        # legend
        legend.position = c(0.12, 0.85), # (right/left, top/bottom)
        legend.box.background = element_rect(color = "black")
        ) 
```

*Approach 2* Shortcut using `stat_summary()`; benefits: (a) no need to create a second df with summary stats, (b) only need to set overall aesthetics once in the ggplot() function
```{r}
##############################
# plot
##############################

# plot raw data
ggplot(data = na.omit(penguins), aes(x = species, y = body_mass_g, shape = sex)) +
  geom_point(alpha = 0.2, position = position_jitterdodge(dodge.width = 0.8)) + 
  
  # calculate and layer on summary stats (mean +/- sd)
  stat_summary(mapping = aes(color = species),
               fun = "mean", geom = "point", size = 4,
               position = position_dodge(width = 0.8)) +
  stat_summary(mapping = aes(color = species),
               fun = "mean", geom = "errorbar", size = 1, width = 0.2,
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               position = position_dodge(width = 0.8)) +
  
  # manually adjust colors and shapes
  scale_color_manual(values = c("darkorange", "purple", "cyan4"), guide = FALSE) + # set `guide = FALSE` to omit from legend
  scale_shape_manual(values = c(15, 16), name = "Sex") + # set `name =` to change legend title
  
  # add nicer axis labels + title + caption
  labs(x = "Penguin Species", 
       y = "Body Mass (g)",
       title = "Body weights of female vs. male adelie, chinstrap, and gentoo penguins",
       subtitle = "Colored points represent mean body mass (± SD)",
       caption = "Data Source: Dr. Kristen Gorman, LTER Palmer Station") +
  
  ########################
  # theme option 1 (basic ggplot)
  ########################
  theme_classic() + # built-in theme
  theme(
        # text
        plot.title = element_text(hjust = 0, size = 14),
        axis.text = element_text(color = "black", size = 10),
        axis.title = element_text(color = "black", size = 14),
        plot.caption = element_text(size = 7, hjust = 1, color = "gray", face = "italic"),

        # plot
        panel.border = element_rect(color = "black", size = 0.7, fill = NA),

        # legend
        legend.position = c(0.12, 0.85), # c(right/left, top/bottom)
        legend.box.background = element_rect(color = "black"))

  ########################
  # theme option 2 (a little bit fancier)
  ########################
  # theme_ipsum(grid = FALSE) + # from hrbrthemes package
  # theme(
  #       # text
  #       plot.title = element_markdown(hjust = 0, size = 13),
  #       plot.caption = element_text(size = 7, hjust = 1, color = "gray", face = "italic"),
  # 
  #       # plot
  #       panel.border = element_rect(colour = "black", size = 0.7, fill = NA),
  # 
  #       # legend
  #       legend.position = c(0.12, 0.85), # c(right/left, top/bottom)
  #       legend.box.background = element_rect(color = "black"))

# title for use with ggtext() : "Body weights of female vs. male <b style='color:darkorange'>adelie,</b> <b style='color:purple'>chinstrap,</b> and <b style='color:cyan4'>gentoo</b> penguins"

########################
# Optional: save your plot as a .jpg, .pdf, etc. using ggsave()
########################

# ggsave(plot = SC_simple, here::here("plots", "SC_simple.jpg"), width = 7, height = 7)
```

