---
title: "juliette_prep"
author: "Juliette Verstaen"
date: "7/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

List of resources we might want to get together and give people
- themes https://ggplot2.tidyverse.org/reference/ggtheme.html
- data viz book https://serialmentor.com/dataviz/visualizing-associations.html
- https://www.data-to-viz.com/about.html=]


Workshop:
https://github.com/allisonhorst/palmerpenguins

Looking at the data and seeing what's what
```{r}
# remotes::install_github("allisonhorst/palmerpenguins")
library(palmerpenguins)
library(ggplot2)
library(tidyverse)
# remotes::install_github("wilkelab/ggtext")
library(hrbrthemes)
```

First things first before making any types of graphs is getting to know your data. For me knowing how many NA's there are and the class of each column/variable is the most important. Not knowing these things up front can cause weird things to happen (or things not to happen) so it's always good to know before hand. It's always best to know the structure of your data well enough that you know the general lay out of your plot. This step is skipped a lot because there's no real product that comes out of it and it takes a little bit of time. In the end though it's so worth it!

Basic look into the data
```{r}
penguins <- penguins

head(penguins)
summary(penguins) ## take the chance to addess the NA values in each column
```

Talk about the different classes and how they might mess up plotting (numeric/ character, an NA might mess that up; factor vs character)

Exploring classes
```{r}
# long way:
class(penguins$species) ## factor

# short way (would need a brief description of what lapply is..)
lapply(penguins, class)
## im actually not super sure when numeric versus integer comes into play for graphing... anyone?
```

Color pallette
```{r}
col_pal <- c("#008A82", "#57280F", "#03304A")
```


```{r}

ggplot(na.omit(penguins)) +
   geom_point(aes(x = body_mass_g, y = flipper_length_mm, color = species, shape = island), size = 2) +
   theme_minimal() +
   
   ## Labeling everything
   labs(x = "Body Mass (g)",
        y = "Flipper Length (mm)",
        title = "Exploring Palmer Station Penguins",
        subtitle = "Dimensions and Attributes",
        caption = "Data Source: Dr. Kristen Gorman LTER Palmer Station",
        shape = "Island",
        color = "Species") +
   
   ## Editing the labels aestetics
   theme(plot.title = element_text(color = "black", size = 18, face = "bold", hjust = 0.5),
         plot.subtitle = element_text(color = "gray41", size = 14, face = "bold", hjust = 0.5),
         plot.caption = element_text(color = "gray65", face = "italic"),
         axis.title.x = element_text(size = 15),
         axis.title.y = element_text(size = 15)) +
   
   ## change colors
   scale_colour_manual(values = col_pal, aesthetics = "colour")+
   facet_grid(cols = vars(sex)) 

ggsave("plots/JV_simple.jpg", width = 10, height = 6, dpi=300) 
```

Make the same plot but without a legend!! And with a pretty new theme. 
```{r}

ggplot(na.omit(penguins)) +
   geom_point(aes(x = body_mass_g, y = flipper_length_mm, color = species)) +
   theme_ipsum_rc() +
   
   ## Labeling everything
   labs(x = "Body Mass (g)",
        y = "Flipper Length (mm)",
        ## use markdown style to code the colors into the title
        title = "Dimensions for <b style='color:#008A82'>adelie,</b> <b style='color:#57280F'>chinstrap,</b> and <b style='color:#03304A'>gentoos</b> penguins",
        caption = "Data Source: Dr. Kristen Gorman LTER Palmer Station") +
   
   ## Editing the labels aesthetics
      theme(plot.title = element_markdown(hjust = 0.3),
            plot.caption = element_text(color = "gray65", face = "italic", size = 8),
            legend.position = "none") +

   ## change colors
   scale_colour_manual(values = col_pal)+
   facet_grid(cols = vars(sex)) 

ggsave("plots/JV_prettier.jpg", width = 10, height = 6, dpi=300) 
```






## Bar plot:
- stacked bar plot for each of the three years with total penguins on the y axis and three chunks of colors for the three penguins, also do it by island

```{r}

penguins_sampled <- penguins %>% 
   mutate(island_add = "Island") %>% 
   unite(island_species, c("species", "island"), sep = " on ", remove = FALSE) %>% 
   unite(island_species, c("island_species", "island_add"), sep = " ") %>% 
   mutate(island_species = fct_relevel(island_species, levels = c("Adelie on Biscoe Island", "Gentoo on Biscoe Island", "Adelie on Dream Island", "Chinstrap on Dream Island", "Adelie on Torgersen Island"))) %>% 
   mutate(year = as.character(year))

colors <- c("#B35B91", "#F0C6E7", "#B1611A", "#F5E3C6", "#4EBCD3")


class(penguins_sampled$year)

ggplot(na.omit(penguins_sampled)) +
  stat_count(aes(x=year, fill=island_species), stat="identity", width = 0.6) +
# geom_text(aes(label = 'island'))
   theme_minimal() +
   labs(x = "Year",
        y = "Sample Count",
        title = "Number of Penguins Sampled") + 
   scale_fill_manual(values = colors, aesthetics = "fill") 


```

biscoe (A, G)
dream (s, c)
torgersen (a)

order factor (bis A, bis G, dream s, dream c, torg a)


## scatter with correlation line
- geom point with bill size and weight, correlation line. bring the point home that you should really explore the data. do one with all the penguins and a line, and then one with 3 lines by penguins. will show that one of the penguins just has a smaller bill size and that you can't generalize by all penguins. maybe do one by penguin species and gender
- remember, it can be really easy to make plots and assume because it looks like something is happening or not that that is the case because you can see it! always explore your data and play around with it, and think critically about what what your doing actually means and why


# 008A82, 57280F, 03304A
## do we want to intoduce spcial themes with all the stuffi did above? or too advanced
## I want to capitilize the male and female labels in the grid but i need to do levels(penguins$sex) <- c("Female", "Male") after doing levels(penguins$sex) to check order
## OR I want to get rid of those titles and put in male or female symbols, can use ggtext thoretically to use picutrees instead https://github.com/wilkelab/ggtext


##do we want to remove NAs? or plot them like allison did 

unique(penguins$bill_length_mm)
summary(penguins$bill_length_mm)

 # scale_size(range = c("32.10", "39.23", "44.45", "43.92", "48.50", "59.60")) +



#############







