---
title: "juliette_prep"
author: "Juliette Verstaen"
date: "7/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

List of resources we might want to get together and give people
- themes https://ggplot2.tidyverse.org/reference/ggtheme.html
- data viz book https://serialmentor.com/dataviz/visualizing-associations.html
- https://www.data-to-viz.com/about.html=]


Workshop:
https://github.com/allisonhorst/palmerpenguins

Load packages
```{r load packages}
# remotes::install_github("allisonhorst/palmerpenguins")
library(palmerpenguins)
library(ggplot2)
library(tidyverse)
library(hrbrthemes)
# remotes::install_github("wilkelab/ggtext")
library(ggtext)
```


## Explore your data

```{r explore data}

head(penguins)
summary(penguins) ## take the chance to address the NA values in each column

## checking out classes (the long way)
class(penguins$species) ## factor

# short way 
lapply(penguins, class)

```


## Color palette

Create a vector of colors that we want to use in our plot. I know I want each penguin to have it's own color, so I picked 3 colors.
```{r color palette}

col_pal <- c("darkorange", "purple", "cyan4")

```

## Create your plot

```{r create simple scatter}

ggplot(na.omit(penguins)) +
   
   ## 1. Choose your plot and aesthetics
   geom_point(aes(x = body_mass_g, y = flipper_length_mm, color = species, shape = island), size = 2) +
   
   ## 2. Label all the things
   labs(x = "Body Mass (g)",
        y = "Flipper Length (mm)",
        title = "Exploring Palmer Station Penguins",
        subtitle = "Dimensions and Attributes",
        caption = "Data Source: Dr. Kristen Gorman LTER Palmer Station",
        shape = "Island",
        color = "Species") +
   
   ## 3a. Built in theme
   theme_minimal() + 
   
   ## 3b. Editing the labels aesthetics
   theme(plot.title = element_text(color = "black", size = 18, face = "bold", hjust = 0.5),
         plot.subtitle = element_text(color = "gray41", size = 14, face = "bold", hjust = 0.5),
         plot.caption = element_text(color = "gray65", face = "italic"),
         axis.title.x = element_text(size = 15),
         axis.title.y = element_text(size = 15)) +
   
   ## 3c. Change colors
   scale_colour_manual(values = col_pal) +
   
   ## 3d. Faceting
   #facet_wrap(vars(sex)) 
   facet_grid(cols = vars(sex), rows = vars(island))


#ggsave("plots/JV_simple.jpg", width = 10, height = 6, dpi=300) 

```

Now we are going to make the same plot but with a pretty new theme and  incorporating the legend into the title!!

```{r}

ggplot(na.omit(penguins)) +
 
   ## 1. Choose your plot and aesthetics
   geom_point(aes(x = body_mass_g, y = flipper_length_mm, color = species)) +
   
   ## 2. Label all the things
   labs(x = "Body Mass (g)",
        y = "Flipper Length (mm)",
        
        ## use html/markdown style to code the colors into the title
        ## title : "Dimensions for adelie, chinstrap, and gentoos penguins"
        ## We need to tell ggplot that:  html format text is coming up, what we want that text to look like, what the text says, and that html format text is over
        title = "Dimensions for <b style='color:darkorange'>adelie,</b> <b style='color:purple'>chinstrap,</b> and <b style='color:cyan4'>gentoos</b> penguins",
        
        caption = "Data Source: Dr. Kristen Gorman LTER Palmer Station \nCreated for RLadies SB workshop") +
   
   ## 3a. Choose an 'hrbrthemes' theme
   theme_ipsum_rc() +
   
   ## 3b. Editing the labels aesthetics
   theme(plot.title = element_markdown(hjust = 0.5, size = 20),
         plot.caption = element_text(color = "gray65", face = "italic", size = 8),
         legend.position = "none") +
   
   ## 3c. Change colors
   scale_colour_manual(values = col_pal) +
   
   ## 3d. Faceting
   facet_grid(cols = vars(sex), rows = vars(island)) 

## ggsave("plots/JV_prettier.jpg", width = 10, height = 6, dpi=300) 
```



