---
title: "Data Viz Part 1 - Plotting summary stats using `stat_summary()`"
author: "Sam Csik"
date: "8/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here, we'll walk through plotting a combination of raw data and summary stats together on the same figure. There are two general approaches:

(1) calculate summary stats and save to a new data frame; plot raw data from original data frame and layer on summary stats from the new data frame -- this tends to require **more** code

(2) use the `stat_summary()` function within ggplot to simultaneously calculate and plot summary stats 

We'll recrete the same figure using both approaches.

####**Part 0.** Import libraries (it's okay if you don't have the naniar and ggtext packages right now)
```{r}
library(palmerpenguins) 
library(tidyverse)
library(ggtext)
library(hrbrthemes)
```

####**Part 1.** Longer way: Summarize data for plotting, then construct figure

*NOTE: Normally, you should take a critical look at your data, determine why NAs exist and if they might bias your results...we're skipping that part here and just omitting any NAs for the sake of time. But for when you do, check out the `naniar` package*
```{r}
##############################
# summarize body mass data (mean, max, min)
##############################

bm_summary <- penguins %>% 
  drop_na() %>% 
  group_by(species, sex) %>% 
  summarize(
    mean = mean(body_mass_g),
    sd = sd(body_mass_g)
  )

##############################
# plot
##############################

ggplot() +
  
  # plot raw data
  geom_point(data = na.omit(penguins), aes(x = species, y = body_mass_g, shape = sex), 
             alpha = 0.2, 
             position = position_jitterdodge(dodge.width = 0.8)) +
  
  # layer on summary stats
  geom_errorbar(data = bm_summary, aes(x = species, ymin = mean - sd, ymax = mean + sd, color = species, group = sex), 
                width = 0.1, size = 1,
                position = position_dodge(width = 0.8)) +
  geom_point(data = bm_summary, aes(x = species, y = mean, shape = sex, color = species),
             size = 4,
             position = position_dodge(width = 0.8)) +
  
  # manually adjust colors and shapes (gotta reflect what's up )
  scale_color_manual(values = c("darkorange", "purple", "cyan4"), guide = FALSE) + # set `guide = FALSE` to omit from legend
  scale_shape_manual(values = c(15, 16), name = "Sex") + # set `name =` to change legend title
  
  # add nicer axis labels + caption
  labs(x = "Penguin Species", 
       y = "Body Mass (g)",
       title = "Body weights of female vs. male <b style='color:darkorange'>adelie,</b> <b style='color:purple'>chinstrap,</b> and <b style='color:cyan4'>gentoo</b> penguins",
       subtitle = "Mean (± SD)",
       caption = "Data Source: Dr. Kristen Gorman, LTER Palmer Station") +
  
  # now fun with themes!
  theme_ipsum(grid = FALSE) +
  theme(
        # text
        plot.title = element_markdown(hjust = 0, size = 13),
        # axis.text = element_text(color = "black", size = 10), 
        # axis.title = element_text(size = 14), 
        plot.caption = element_text(size = 7, hjust = 1, color = "gray", face = "italic"),
        # plot
        panel.border = element_rect(colour = "black", size = 0.7, fill = NA), 
        # legend
        legend.position = c(0.12, 0.85), # (right/left, top/bottom)
        legend.box.background = element_rect(color = "black")
        ) 
```

####**Part 2.** Shortcut using `stat_summary()`; benefits: (a) no need to create a second df with summary stats, (b) only need to set overall aesthetics once in the ggplot() function
```{r}
##############################
# plot
##############################

# plot raw data
ggplot(data = na.omit(penguins), aes(x = species, y = ___, shape = ___)) +
  geom_point(alpha = ___, position = position_jitterdodge(dodge.width = ___)) + 
  
  # calculate and layer on summary stats (mean +/- sd)
  stat_summary(mapping = aes(color = ___),
               fun = "___", geom = "___", size = ___,
               position = position_dodge(width = ___)) +
  stat_summary(mapping = aes(color = species),
               fun = "___", geom = "___", size = ___, width = ___,
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               position = position_dodge(width = ___)) +
  
  # manually adjust colors and shapes
  scale_color_manual(values = c("___", "___", "___"), guide = FALSE) + # set `guide = FALSE` to omit from legend
  scale_shape_manual(values = c(___, ___), name = "Sex") + # set `name =` to change legend title
  
  # add nicer axis labels + title + caption
  labs(x = "Penguin Species", 
       y = "___",
       title = "Body weights of female vs. male penguins",
       subtitle = "Mean (± SD)",
       caption = "Data Source: Dr. Kristen Gorman, LTER Palmer Station") +
  
  ########################
  # theme option 1 (basic ggplot)
  ########################
  theme_classic() + # built-in theme
  theme(
        # text
        plot.title = element_text(hjust = ___, size = ___),
        axis.text = element_text(color = "___", size = ___),
        axis.title = element_text(color = "___", size = ___),
        plot.caption = element_text(size = ___, hjust = ___, color = "___", face = "___"),

        # plot
        panel.border = element_rect(color = "___", size = ___, fill = NA),

        # legend
        legend.position = c(___, ___), # c(right/left, top/bottom)
        legend.box.background = element_rect(color = "___"))

  ########################
  # theme option 2 (a little bit fancier)
  ########################
  # theme_ipsum(grid = FALSE) + # from hrbrthemes package
  # theme(
  #       # text
  #       plot.title = element_markdown(hjust = ___, size = ___),
  #       plot.caption = element_text(size = ___, hjust = ___, color = "___", face = "___"), 
  #       
  #       # plot
  #       panel.border = element_rect(colour = "___", size = ___, fill = NA), 
  #       
  #       # legend
  #       legend.position = c(___, ___), # c(right/left, top/bottom)
  #       legend.box.background = element_rect(color = "___"))

# title for use with ggtext() : "Body weights of female vs. male <b style='color:darkorange'>adelie,</b> <b style='color:purple'>chinstrap,</b> and <b style='color:cyan4'>gentoo</b> penguins"
```
